---
title: "Summary of the autosomal SNP data of Ritidian Beach Cave, Guam"
author: "Alex Hübner"
date: "`r format(Sys.time(), '%b %d, %Y')`"
output:
    pdf_document:
        fig_caption: yes
bibliography: 191208_autosomalData.bib
---

```{r libraries, echo=F}
library(knitr)
library(data.table)
library(tidyverse)
library(pander)
opts_chunk$set(echo=F, warning=F, message=F)
```
```{r load_data}
libraries <- fread("../05-results/SUM_libraries.csv")
snps <- fread("../05-results/SUM_SNPs.csv")
```

From two samples from Ritidian Beach Cave, Guam, DNA was extracted multiple times and Illumina
sequencing libraries were generated (Table 1). The produced sequencing libraries were initially
screened using shotgun sequencing before capture-enrichment for both mtDNA and autosomal SNPs from
the 1240K array [@Haak2015, @Fu2015] was performed and the enriched libraries paired-end sequenced
across multiple sequencing runs.

```{r add_metadata, results = "asis"}
metadata <- tibble(sampleID = c("SP4210", "SP4211"),
                   sample = c("RBC1", "RBC2"),
                   description = c("Ritidian Beach Cave, Guam",
                                   "Ritidian Beach Cave, Guam")) %>%
            left_join(libraries %>%
                      group_by(sampleID) %>%
                      summarise(`no. of libraries` = length(unique(libraryID))),
                      by = "sampleID")
pandoc.table(metadata %>%
             select(name = sample, sampleID, `no. of libraries`),
             split.table = Inf,
             caption = "Number of produced Illumina sequencing libraries for Ritidian Beach Cave samples.")
```

The sequencing data was processed by the following steps. First, Illumina sequencing adapters were
removed using *leehom* [@Renaud2014] and overlapping read pairs were merged. The sequencing data
de-multiplexed by assigning reads only to a sample when there were no mismatches between the
observed and expected double-index combination of seven basepair length. The sequencing data was
subsequently aligned against the human reference genome *hg19* from the 1000Genomes projects with
decoy sequences
(http://ftp.1000genomes.ebi.ac.uk/vol1/ftp/technical/reference/GRCh38_reference_genome/GRCh38_full_analysis_set_plus_decoy_hla.fa)
using BWA *aln* [@Li2009] with the settings "-n 0.01 -o 2 -l 16500" to allow for a higher number of
mismatches and switch off the seeding algorithm. We filtered the aligned sequencing data and only
kept reads for which the paired reads could be merged, had a minimum read length of 35 bp, and a
minimal mapping quality of 25. Finally, duplicated reads were identified and removed using DeDup
[@Peltzer2016].

```{r filtering_summary, results = "asis"}
filtering_stats <- libraries %>%
                   filter(array == "MT") %>%
                   group_by(sampleID, libraryID) %>%
                   summarise(`raw reads` = sum(`raw reads`),
                             `reads w/ L35` = sum(`reads w/ L35`),
                             `reads with L35 [%]` = round(`reads w/ L35` / `raw reads` * 100, 2),
                             `reads with MQ25 [%]` = round(sum(`reads w/ MQ25`) / `raw reads` * 100, 2),
                             `final reads` = sum(`unique reads`),
                             `duplication rate` = mean(`duplication rate`)) %>%
                   ungroup() %>%
                   mutate(sample = c("RBC1", rep("", 8), "RBC2", rep("", 6))) %>%
                   select(sample, `libraryID`:`duplication rate`)
pandoc.table(filtering_stats,
             split.table = Inf,
             split.cell = 40,
             caption = "Summary of the filtering process summarised by library. The percentages of
             the reads with length >= 35 bp [L35] and the mapping quality >= 25 [MQ25] are based on
             the raw reads.")
```


# References
